### 2016-12-15 paper/5_evaluation.tex
Spell checking

The results depicted in \cref{fig:mmNodesRuntime} indicate for both execution engines decent scale-out behavior.
For Spark we observe an almost horizontal line for number of cores $\le 32$.
For larger number of cores, the scaling degrades which is probably caused by spilling.
The same applies to Flink.
However, we can observe that the system has to start spilling data earlier.
Thus, Flink's scale-out behaviour is not as good as Spark's with respect to matrix multiplication.

@@

The graphs in \cref{fig:pageRankResults} show that the specialized PageRank algorithm runs faster than Gilbert's versions of PageRank.
Furthermore, the specialized algorithms show a better scalability.
The speedup of the specialized algorithms with respect to Gilbert's implementations for this experiment is shown in \cref{fig:pageRankSpeedup}.
For $n\le 50000$, the hand-tuned algorithms running on Spark and Flink show a similar speedup.
The Flink and Spark version achieve a speedup of approximately $13$ and $12$ for $n=50000$, respectively.
However, for $n = 100000$ we can observe a speedup of $76$ for the Flink specialized implementation whereas Spark reaches a speedup of $14$.
Gilbert's performance degradation with Flink is likley caused by earlier data spilling.

>>>>

The results depicted in \cref{fig:mmNodesRuntime} indicate for both execution engines decent scale-out behavior.
For Spark we observe an almost horizontal line for number of cores $\le 32$.
For larger number of cores, the scaling degrades which is probably caused by spilling.
The same applies to Flink.
However, we can observe that the system has to start spilling data earlier.
Thus, Flink's scale-out behavior is not as good as Spark's with respect to matrix multiplication.

@@

The graphs in \cref{fig:pageRankResults} show that the specialized PageRank algorithm runs faster than Gilbert's versions of PageRank.
Furthermore, the specialized algorithms show a better scalability.
The speedup of the specialized algorithms with respect to Gilbert's implementations for this experiment is shown in \cref{fig:pageRankSpeedup}.
For $n\le 50000$, the hand-tuned algorithms running on Spark and Flink show a similar speedup.
The Flink and Spark version achieve a speedup of approximately $13$ and $12$ for $n=50000$, respectively.
However, for $n = 100000$ we can observe a speedup of $76$ for the Flink specialized implementation whereas Spark reaches a speedup of $14$.
Gilbert's performance degradation with Flink is likely caused by earlier data spilling.

### 2016-12-15 paper/6_related_work.tex
Spell checking

{\em SystemML}~\cite{ghoting:2011a,Boehm2014,Boehm2014Hybrid,Schelter2015,Elgohary2016} aims to make machine learning algorithms run on massive datasets without burdening the user with low-level implementation details and tedious hand-tuning. 
Therefore, it provides an R-like declarative high-level language, called Declarative Machine learning Language (\emph{DML}), and compiles and optimizes the resulting programs to distributed dataflow systems. 
The linear algebra operations are translated into a directed acyclic graph of high-level operators. 
This abstract representation allows to apply several optimizations such as algebraic rewrites, choice of internal matrix representation and cost-based selection of the best execution plan. 
Afterwards these plans are translated to low-level operators and executed either in the driver memory or distributedly. 
SystemML is one of the main inspirations for Gilbert. 
While it originally only supported MapReduce as runtime, it has recently also moved to supporting more advanced dataflow systems. 
Gilbert differs from SystemML by its fixpoint operator and by leveraging the general optimizers of the underlying system (e.g., the Flink optimizer for matrix multiplication optimization, see~Section~\ref{sec:LinearAlgebraOperations}).

@@

{\em Ricardo}~\cite{das:2010a} uses existing technologies to implement a scalable system for statistical analysis: it executes the data shipment via Hadoop and runs the actual analytic computation by R. 
It integrates these systems via Jaql~\cite{beyer:2011a}, a declarative high-level language for data-processing on Hadoop. 
Thereby, the user can initiate the distribution, transformation and aggregation of data within Hadoop. 
Furthermore, the system supports to run R code on the worker nodes as data transformations. 
However, in this setting the user still has to explicitly specify the data-flow and the data distribution, which requires a substantial understanding of MapReduce's principles. 
{\em RHIPE}~\cite{guha:s2012a} also tries to extend R to scale to large data sets using Hadoop. 
RHIPE follows an approach called divide and recombine. 
The idea is to split the examined data up into several chunks so that they fit in the memory of a single computer. 
Next a collection of analytic methods is applied to each chunk without communicating with any other computer. 
After all chunks are evaluated, the results will be recombined in an aggregation step to create the overall analytic result. 
However, this strict execution pipeline constrains the analysis process considerably. 
A system which integrates more seamlessly into the R ecosystem is {\em pR} (parallel R)~\cite{samatova:2009a}. 
The goal is to let statisticians compute large-scale data without having to learn a new system. 
pR achieves its goal by providing a set of specialized libraries which offer parallelized versions of different algorithms using MPI.
However, MPI does not integrate well with an evironment where clusters are shared and usually execute several jobs concurrently. 
Furthermore, it lacks important properties such as fault-tolerance, elasticity and reliability. 
Another related system for R is {\em rmpi}~\cite{rmpi}. 
As this is only a wrapper for the respective MPI calls, it also suffers from the aforementioned problems. 
Furthermore, there is the SparkR~\cite{Venkataraman2016} project which aims to integrate Spark's API and SQL processing capabilities into R.

>>>>

{\em SystemML}~\cite{ghoting:2011a,Boehm2014,Boehm2014Hybrid,Schelter2015,Elgohary2016} aims to make machine learning algorithms run on massive datasets without burdening the user with low-level implementation details and tedious hand-tuning. 
Therefore, it provides an R-like declarative high-level language, called Declarative Machine learning Language (\emph{DML}), and compiles and optimizes the resulting programs to distributed dataflow systems. 
The linear algebra operations are translated into a directed acyclic graph of high-level operators. 
This abstract representation allows to apply several optimizations such as algebraic rewrites, choice of internal matrix representation and cost-based selection of the best execution plan. 
Afterwards these plans are translated to low-level operators and executed either in the driver memory or in parallel. 
SystemML is one of the main inspirations for Gilbert. 
While it originally only supported MapReduce as runtime, it has recently also moved to supporting more advanced dataflow systems. 
Gilbert differs from SystemML by its fixpoint operator and by leveraging the general optimizers of the underlying system (e.g., the Flink optimizer for matrix multiplication optimization, see~Section~\ref{sec:LinearAlgebraOperations}).

@@

{\em Ricardo}~\cite{das:2010a} uses existing technologies to implement a scalable system for statistical analysis: it executes the data shipment via Hadoop and runs the actual analytic computation by R. 
It integrates these systems via Jaql~\cite{beyer:2011a}, a declarative high-level language for data-processing on Hadoop. 
Thereby, the user can initiate the distribution, transformation and aggregation of data within Hadoop. 
Furthermore, the system supports to run R code on the worker nodes as data transformations. 
However, in this setting the user still has to explicitly specify the data-flow and the data distribution, which requires a substantial understanding of MapReduce's principles. 
{\em RHIPE}~\cite{guha:s2012a} also tries to extend R to scale to large data sets using Hadoop. 
RHIPE follows an approach called divide and recombine. 
The idea is to split the examined data up into several chunks so that they fit in the memory of a single computer. 
Next a collection of analytic methods is applied to each chunk without communicating with any other computer. 
After all chunks are evaluated, the results will be recombined in an aggregation step to create the overall analytic result. 
However, this strict execution pipeline constrains the analysis process considerably. 
A system which integrates more seamlessly into the R ecosystem is {\em pR} (parallel R)~\cite{samatova:2009a}. 
The goal is to let statisticians compute large-scale data without having to learn a new system. 
pR achieves its goal by providing a set of specialized libraries which offer parallelized versions of different algorithms using MPI.
However, MPI does not integrate well with an environment where clusters are shared and usually execute several jobs concurrently. 
Furthermore, it lacks important properties such as fault-tolerance, elasticity and reliability. 
Another related system for R is {\em rmpi}~\cite{rmpi}. 
As this is only a wrapper for the respective MPI calls, it also suffers from the aforementioned problems. 
Furthermore, there is the SparkR~\cite{Venkataraman2016} project which aims to integrate Spark's API and SQL processing capabilities into R.

### 2016-12-15 paper/7_conclusion.tex
Spell checking

>>>>

\textbf{Acknowledgments.}
This work has been partially supported through grants by the German Federal Ministry for Education and Research for the project Berlin Big Data Center (funding mark 01IS14013A) and by the German Federal Ministry of Economic Affairs and Energy for the project Smart Data Web (funding mark 1MD15010A).
